What is an Ancestral Recombination Graph?

### 2022-03-10 paper.tex
Describe implications of fig 3 in 3 parts
As discussed with @jeromekelleher

In this paper we [turn into proper round up paragraph later]
\begin{enumerate}
\item Discuss the classical Ancestral Recombination Graph formulation,
and show how it derives from (and is limited by) the coalescent with
recombination.
\item Suggest a definition of ARGs that is free of these limitations,
where the formulation is derived from a (normally diploid) pedigree rather than from
coalescent approximations, and where the genetic
material which is transmitted is described by annotations attached to
edges rather than nodes.
\item Suggest a classification of ARG nodes that helps us to understand
the properties of the structures inferred by different methods, and illustrate
with some examples.
% Perhaps we should also mention something about simplification and unknowable
% nodes, e.g. "show how some of the nodes are not required for useful interpretation
% of ancestry".
\item Throughout, we consider ARGs defined for a single chromosome comprising
of a fixed number of discrete sites. Our argument applies equally well to ARGs
defined for several non-homologous chromosomes \citep{fearnhead2003ancestral,
koskela2019robust} and to idealised, continuous chromosomes, but addressing
all these cases simultaneously would result in needlessly bulky notation.
\end{enumerate}

@@

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11, 1.2);
	\draw[color=gray, dashed] (7.5, 1.7) -- (11, 1.7);
	\draw[color=gray, dashed] (7.5, 2.4) -- (11, 2.4);
	\draw[color=gray, dashed] (7.5, 3.1) -- (11, 3.1);
	\draw[color=gray, dashed] (7.5, 3.8) -- (11, 3.8);
	\draw[color=gray, dashed] (7.5, 4.5) -- (11, 4.5);
	\draw[color=gray, dashed] (7.5, 5.4) -- (11, 5.4);
	\draw[color=gray, dashed] (7.5, 6) -- (11, 6);
	\draw[color=gray, dashed] (7.5, 6.6) -- (11, 6.6);
	\draw[color=gray, dashed] (7.5, 7.2) -- (11, 7.2);
\end{tikzpicture}
\end{tabular}
}
\caption{(a)
A realisation of the graph traversed by Hudson's algorithm started from a
sample of three chromosomes of length $m$ at time $t = 0$, and
propagated until time $T$. The MRCA on the genetic interval $[v, w)$ is reached
at event B, while that on $[0, v)$ is reached at event C.
The non-ancestral segment $[v, w)$ above
A contributes to the rate of effective recombinations because it
is trapped between ancestral segments. The two columns titled CA and RE
are the respective rates of mergers and recombinations when
the recombination rate is $\rho$.
(b) A corresponding realisation of a big ARG, which augments Hudson's algorithm
by tracking nonancestral lineages. The result is a simpler state space and
dynamics, at the cost of extra nodes and edges, highlighted in red, which do
not affect the local tree at any site.}
\label{hudson_vs_bigARG}
\end{figure}

@@

The ordering of edges is a vital element of the EARG encoding.
This is because there is a fundamental ambiguity involved
in associating recombination breakpoints with \emph{nodes}
rather than edges,
and the only way we can break this symmetry is to assign
some meaning to the order in which the parents are listed
(\cite{griffiths1997ancestral} refer to the ``left'' and ``right''
parents of a recombination event). This is usually described
in terms of the process in which we extract the local trees
along the genome from an EARG.
To recover
the tree at genome coordinate $x$,
we traverse the graph upwards from the leaves. At a particular
node $u$, if it has one parent we are at a common ancestor
node and we simply follow that parent. If we are at a
recombination node, $u$ has two parents: if
$x$ is less than the breakpoint $\sigma(u)$ we follow
the edge to the first parent, and otherwise follow the edge
to the second parent.
This ordering requirement, which is straightforward
to describe, has some practical drawbacks. For example,
% Surely the meaning is clear from the context here and we don't need
% qualify the meaning of "ARG"
simulating an ARG in this representation is
complicated by the fact that the
first event to occur may hit the lineage carrying the ancestry
to the right of the breakpoint rather
than the left (and so we cannot emit edges as they are generated).
Such issues can be worked around, of course,
but depending on the ordering of otherwise indistinguishable
objects is generally problematic. In practice, several
methods
explicitly associate information with the outbound edges
of a recombination event
to resolve the problem~\citep{lyngso2005minimum,ignatieva2021kwarg}.

@@

To distinguish from the classical event-based encodings,
we call this encoding of an ARG data
structure a ``Genome ARG'', or GARG.
Fig~\ref{fig-arg-in-pedigree} illustrates the basic elements of a GARG.
For concreteness, Fig.~\ref{fig-arg-in-pedigree}a shows the genomes
embedded in a inbred diploid pedigree, in which each
individual carries two genomes (each individual, labelled $i_1$ to $i_8$,
therefore has a paternal genome on the right and a maternal one on
the left, although in principle there is no requirement for a specific mating
system or ploidy). A genome is inherited from one of the individual's parents,
and may be the recombined product of that parent's two genomes.
For example, individual $i_1$ has two genomes \textsf{A} and \textsf{B},
inherited from parents $i_3$ and $i_4$. Genome \textsf{A} is the product of
recombining $i_3$'s two genomes \textsf{E} and \textsf{F} at position 2,
whereas \textsf{B} was inherited directly from $i_4$'s genome \textsf{G} without
recombination. Fig.~\ref{fig-arg-in-pedigree}b illustrates that the patterns of
transmission between genomes can be followed without necessarily knowing
how they are combined within individuals. This ``genome-perspective''
simply depicts the ancestry of the sample genomes $A$--$D$. The graph
topology along with the genome coordinate annotations on the
edges then uniquely defines the trees at every position
along the genome (Fig.~\ref{fig-arg-in-pedigree}c).

The focus on modelling genomes rather than events has some interesting
and important consequences. Firstly, the graph may contain nodes that
do not correspond to any particular event in the ancestry of the samples.
For example, node \textsf{H} is the direct ancestor of \textsf{D} in
Fig~\ref{fig-arg-in-pedigree}, and is not the product of either a
coalescence or a recombination. Such nodes would usually be removed
from the graph so that \textsf{D} descends directly from \textsf{K} (see
section XXX) but there is no necessity for this from a representational
perspective. This ability can be useful in practice and could be required if,
% Do we need to spell this out?
for example, we have a deeply sequenced pedigree of
haplodiploids. % e.g. Honeybees? Be great to cite an example!

Secondly, the way that we model recombination is subtly different.
Rather than having a single node that represents the \emph{event},
there are three nodes (representing the recombinant offspring and
two parental genomes) and two edges with associated genome coordinate
intervals (representing the precise inheritance patterns).
For example, in Fig~\ref{fig-arg-in-pedigree}, node \textsf{C} is the
recombinant descendent of \textsf{E} and \textsf{F}, inheriting positions
0 to 7 from $E$ and genome positions 7 to 10 from $F$. Any particular
node can be both a recombinant (it is the recombined product of
its parent's genomes) and a common ancestor (the ancestry from
multiple samples meet at this node; see the next section).
% Here was can talk about representing the recombination by a
% recombinant and 2 parent "recombination nodes" rather than
% a single "recombination node" as in the normal ARG representation.
% We may also wish to note that in this example the recombination nodes
% are shared between siblings, and can also be coalescent nodes
% this ties in with stuff discussed 3 paras below.
[More stuff here: clarify why we assign coordinates to edges
rather than nodes. Basically lay out all the stuff that we can
talk about \emph{without} going in to the details of
ancestry resolution.]

@@

% Asking the reader to flick back to figure 2 here, not great.
The differences between EARGs
and GARGs are subtle but important, and illustrated in
Fig.~\ref{fig-arg-data-structure}, which shows a simple ARG
with three samples and one recombination.
Fig.~\ref{fig-arg-data-structure}A shows this topology as a
classical Griffiths-like EARG, and
Fig.~\ref{fig-arg-data-structure}B shows it as a GARG.
The two are almost identical, except we model the recombination
event via the two parental genomes in nodes $3$ and $4$ and explicitly
store the genomic interval that the child inherited from the
parent with every edge.
This approach may seem wasteful since we must store an extra
node for each recombination and genomic interval information
with every edge (even though many simply span the full genome).
However, asymptotically the amount of information stored is the
same, and as we see in the next section, keeping interval
information for all edges enables useful generalisations.
There is also much simplicity to be gained by having
a single type of node and by treating all edges in the same way.

@@

% AI: have not mentioned any implicit or un-rooted networks (split, median, haplotype, reticulograms, etc).
% Also: are we not defining ARGs as "binary networks"? Might be good to specify this here.

@@

\begin{figure}
\centering
\vspace{5em}
\includegraphics[width=\linewidth]{illustrations/simplification}
\caption{\label{fig-simplification}
ARG simplification. (a) An ARG and equivalent local trees simulated
from a diploid Wright-Fisher
model. Note the complex full-sib relationship between B and C; this
could not be directly represented in a Griffiths-like EARG encoding
(see section XXX).
(b) Simplified to remove all
% trying out this terminology - can define in the text
1-connected graph components (e.g., diamonds).
(c) Remove nodes that are unary in all local trees.
(d) Rewrite edges to skip any unary nodes in local trees.
}
\end{figure}

% Q: would it help to talk about "tree vertices" here or would it
% just be more confusing? I worry that people would think that
% they are two different things then, and it's really important
% that they get that a node-is-a-node
The ideas of ``meaningful'' local tree topologies revolve around the
presence of ``unary'' nodes: a node in the tree that has exactly
one child. For example, Fig.~\ref{fig-simplification} shows a set of
four simulated gARGs and the local trees that can be constructed from
them (for simplicity, edge annotations are not shown)
Consider the right-most tree in Fig.~\ref{fig-simplification}a
and the path above node $D$. Before $D$ meets its common ancestor
over the interval $(93, 100]$ with the other samples it passes through
nodes $H, K, N, R$ and $S$, reflecting the corresponding
traversal through the ARG (see section XXX). Such unary nodes
in a local tree are not usually considered significant,
because they provide very little information and are not
detectable by information from the tree. % weak
The nodes in a tree usually mark the points where samples find a
\emph{common} ancestor, rather than focusing on specific
ancestors. For most purposes, we would consider an edge
directly joining sample $D$ to the node $T$ to be more
useful and informative. Simplification is the process of removing
these uninformative and (for most computational purposes)
inefficient unary paths.

The first type of simplification that we can perform is to remove
graph topology that is invisible to the samples. The most well
known example of such topology is a so-called
``diamond''~\citep{rasmussen2014genome}
in which the two parent nodes of a recombination immediately
join again into a common ancestor (e.g. $K,N,O$ and $R$ in
Fig.~\ref{fig-simplification}a).
Unless we are specifically
interested in the recombination event or ancestral genomes,
there is no information in this topology and the diamond can b
e replaced by a single edge. More generally, any
subgraph which is singly-connected in both the leafward and
rootward direction (a ``super-diamond'') is non-identifiable and can be
replaced by a single edge. This definition includes the case
of a single node that has one inbound and one outbound edge.
Fig.~\ref{fig-simplification}b shows the result of this type of
graph topology simplification.

Simplifying away diamonds will remove many unary nodes from the
local trees, there can still be nodes that are unary in all
of the local trees. For example, node $H$ is one of the
parents of the recombinant node $D$ is still present in
Fig.~\ref{fig-simplification}b after diamond removal, but
it is unary in all of the local trees since it has an
in-degree of $1$.
Fig.~\ref{fig-simplification}c shows the resulting graph
after removing these nodes.

The remaining nodes are MRCAs of some subset of the samples
at \emph{some} positions along the genome. We still have
some unary nodes in the local trees, but these nodes will
correspond to a coalescence in at least one other
local tree. For example, node $E$ is unary in the first tree
of Fig.~\ref{fig-simplification}a, but is either binary
or ternary in the other trees (recall this is a Wright-Fisher
simulation).

The final level of simplification is to remove \emph{all}
unary nodes from the local trees (Fig.~\ref{fig-simplification}d).
[And continue]

@@

Due to the computational expense of sampling ``little ARGs", state-of-the-art
Monte Carlo methods typically rely on cheaper, approximate models.
The most prominent examples are \texttt{ARGWeaver} \citep{rasmussen2014genome}
and \texttt{Arbores} \citep{heine2018bridging}. The former also utilises a time
discretisation approximation, but scales to dozens of mammal-like genomes.
The \texttt{ClonalOrigin} method \citep{didelot2010inference,
medina2020speeding} uses the \texttt{ClonalFrame}
model and data structure, though it is not as scalable as \texttt{ARGWeaver}.

>>>>

In this paper we [turn into proper round up paragraph later]
\begin{enumerate}
\item Discuss the classical Ancestral Recombination Graph formulation,
and show how it derives from (and is limited by) the coalescent with
recombination.
\item Suggest a definition of ARGs that is free of these limitations,
where the formulation is derived from a (normally diploid) pedigree rather than from
coalescent approximations, and where the genetic
material which is transmitted is described by annotations attached to
edges rather than nodes.
\item Suggest a classification of ARG nodes that helps us to understand
the properties of the structures inferred by different methods, and illustrate
with some examples.
% Perhaps we should also mention something about simplification and unknowable
% nodes, e.g. ``show how some of the nodes are not required for useful interpretation
% of ancestry''.
\item Throughout, we consider ARGs defined for a single chromosome comprising
of a fixed number of discrete sites. Our argument applies equally well to ARGs
defined for several non-homologous chromosomes \citep{fearnhead2003ancestral,
koskela2019robust} and to idealised, continuous chromosomes, but addressing
all these cases simultaneously would result in needlessly bulky notation.
\end{enumerate}

@@

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11, 1.2);
	\draw[color=gray, dashed] (7.5, 1.7) -- (11, 1.7);
	\draw[color=gray, dashed] (7.5, 2.4) -- (11, 2.4);
	\draw[color=gray, dashed] (7.5, 3.1) -- (11, 3.1);
	\draw[color=gray, dashed] (7.5, 3.8) -- (11, 3.8);
	\draw[color=gray, dashed] (7.5, 4.5) -- (11, 4.5);
	\draw[color=gray, dashed] (7.5, 5.4) -- (11, 5.4);
	\draw[color=gray, dashed] (7.5, 6) -- (11, 6);
	\draw[color=gray, dashed] (7.5, 6.6) -- (11, 6.6);
	\draw[color=gray, dashed] (7.5, 7.2) -- (11, 7.2);
\end{tikzpicture}
\end{tabular}
}
\caption{(A)
A realisation of the graph traversed by Hudson's algorithm started from a
sample of three chromosomes of length $m$ at time $t = 0$, and
propagated until time $T$. The MRCA on the genetic interval $[v, w)$ is reached
at event B, while that on $[0, v)$ is reached at event C.
The non-ancestral segment $[v, w)$ above
A contributes to the rate of effective recombinations because it
is trapped between ancestral segments. The two columns titled CA and RE
are the respective rates of mergers and recombinations when
the recombination rate is $\rho$.
(B) A corresponding realisation of a big ARG, which augments Hudson's algorithm
by tracking nonancestral lineages. The result is a simpler state space and
dynamics, at the cost of extra nodes and edges, highlighted in red, which do
not affect the local tree at any site.}
\label{hudson_vs_bigARG}
\end{figure}

@@

The ordering of edges is a vital element of the EARG encoding.
This is because there is a fundamental ambiguity involved
in associating recombination breakpoints with \emph{nodes}
rather than edges,
and the only way we can break this symmetry is to assign
some meaning to the order in which the parents are listed
(\cite{griffiths1997ancestral} refer to the ``left'' and ``right''
parents of a recombination event). This is usually described
in terms of the process in which we extract the local trees
along the genome from an EARG.
To recover
the tree at genome coordinate $x$,
we traverse the graph upwards from the leaves. At a particular
node $u$, if it has one parent we are at a common ancestor
node and we simply follow that parent. If we are at a
recombination node, $u$ has two parents: if
$x$ is less than the breakpoint $\sigma(u)$ we follow
the edge to the first parent, and otherwise follow the edge
to the second parent.
This ordering requirement, which is straightforward
to describe, has some practical drawbacks. For example,
% Surely the meaning is clear from the context here and we don't need
% qualify the meaning of ``ARG''
simulating an ARG in this representation is
complicated by the fact that the
first event to occur may hit the lineage carrying the ancestry
to the right of the breakpoint rather
than the left (and so we cannot emit edges as they are generated).
Such issues can be worked around, of course,
but depending on the ordering of otherwise indistinguishable
objects is generally problematic. In practice, several
methods
explicitly associate information with the outbound edges
of a recombination event
to resolve the problem~\citep{lyngso2005minimum,ignatieva2021kwarg}.

@@

To distinguish from the classical event-based encodings,
we call this encoding of an ARG data
structure a ``Genome ARG'', or GARG.
Fig~\ref{fig-arg-in-pedigree} illustrates the basic elements of a GARG.
Fig.~\ref{fig-arg-in-pedigree}a shows the genomes
embedded in an inbred diploid pedigree, in which individuals occur at
discrete generations and carry two genomes. Each individual (labelled
$i_1$ to $i_8$) therefore has a paternal genome (on the right) and a maternal one
(on the left), although the pedigree is only shown for concreteness, and 
there is no inherent limitation on mating system, ploidy, or age structure.
A genome is inherited from one of the individual's parents,
and may be the recombined product of that parent's two genomes.
For example, individual $i_1$ has two genomes \textsf{A} and \textsf{B},
inherited from parents $i_3$ and $i_4$. Genome \textsf{A} is the product of
recombining $i_3$'s two genomes \textsf{E} and \textsf{F} at position 2,
whereas \textsf{B} was inherited directly from $i_4$'s genome \textsf{G} without
recombination. Fig.~\ref{fig-arg-in-pedigree}b shows the bare gARG, illustrating
that the patterns of
transmission between genomes can be followed without necessarily knowing
how they are combined within individuals. This ``genome-perspective''
simply depicts the ancestry of the sample genomes $A$--$D$. The graph
topology along with the genome coordinate annotations on the
edges then uniquely defines the trees at every position
along the genome (Fig.~\ref{fig-arg-in-pedigree}c).

@@

The focus on genomes rather than events has some interesting
and important ramifications. Firstly, a gARG may contain nodes that
do not correspond to any particular event in the ancestry of the samples.
For example, node \textsf{H} is the direct ancestor of \textsf{D} in
Fig~\ref{fig-arg-in-pedigree}, and is not the product of either a
coalescence or a recombination. Such nodes would usually be removed
from the graph so that \textsf{D} descends directly from \textsf{K} (see
section XXX) but there is no necessity for this from a representational
perspective. This ability can be useful in practice and could be required if,
% Do we need to spell this out?
for example, we have a deeply sequenced pedigree of
haplodiploids. % e.g. Honeybees? Be great to cite an example!

Secondly, unlike in an eARG, nodes in a gARG are not classified into different
types. Indeed a single node can have both multiple children
and multiple parents. This would be the case in Fig~\ref{fig-arg-in-pedigree} if,
for instance, node \textsf{A} were to have two children.
% It is a shame that this is not actually shown in fig 3, but I don't think
% we can do this unless we have 5 generations
Moreover, a single node can have more than 2 children, and (as we shall see in
section XXX), more than 2 parents. Another way to state this is that transitions
between nodes in a gARG can encapsulate multiple events in an eARG.

The final point deals with the way in which recombination is represented, and is
closely linked to which genomes are chosen as the nodes in a gARG.
We could potentially choose genomes at any point in the life cycle: for instance,
we could chose to represent gametes as ARG nodes. However, the most intuitive
choice is where genomes represent diploid individuals in each generation, as in
Fig~\ref{fig-arg-in-pedigree}a. In this case recombination
is not represented by a single ``recombination node'' (as in an eARG), but with
two parent genomes between which recombination takes place (e.g.
nodes \textsf{E} and \textsf{F}), and a
single resulting ``recombinant'' genome (e.g. \textsf{C}). The
edges that link this recombinant with its two parents carry the information
concerning breakpoint location that would normally be present on the
``recombination nod'' in an eARG. For example, node \textsf{C} in
Fig~\ref{fig-arg-in-pedigree} inherits genome positions 0 to 7 from \textsf{E}
and positions 7 to 10 from \textsf{F}, indicating that the breakpoint is at position 7.

This method for representing recombinations may seem wasteful,
since more than one node is required for each recombination,
and genomic interval information must be stored on every edge.
However, asymptotically the amount of information stored is the
same, and there is much simplicity to be gained by having
a single type of node and by treating all edges in the same way.
More importantly --- as we see in the next section --- we gain the
ability to simplify the ARG by removing nodes without affecting the
local tree topologies, and also the ability to record only the
regions of genome relevant to the final, sample genomes. This last
feature turns out to be the key to efficient processing of the ARG.

[Possibly more stuff here: clarify why we assign coordinates to edges
rather than nodes. Basically lay out all the stuff that we can
talk about \emph{without} going in to the details of
ancestry resolution.]

@@

% AI: have not mentioned any implicit or un-rooted networks (split, median, haplotype, reticulograms, etc).
% Also: are we not defining ARGs as ``binary networks''? Might be good to specify this here.

@@

\begin{figure}
\centering
\vspace{5em}
\includegraphics[width=\linewidth]{illustrations/simplification}
\caption{\label{fig-simplification}
ARG simplification. (A) An ARG simulated from a diploid Wright-Fisher
model. Note the complex full-sib relationship between B and C; this
could not be directly represented in a Griffiths-like EARG encoding
(see section XXX).
(B) Simplified to remove all
% trying out this terminology - can define in the text
1-connected graph components (e.g., diamonds).
(C) Remove nodes that are unary in all local trees.
(D) Rewrite edges to skip any unary nodes in local trees.
(E) The local trees for ARG (A).
(F) The local trees for ARG (D).
}
\end{figure}

% Q: would it help to talk about ``tree vertice'' here or would it
% just be more confusing? I worry that people would think that
% they are two different things then, and it's really important
% that they get that a node-is-a-node
The ideas of ``meaningful'' local tree topologies revolve around the
presence of ``unary'' nodes: a node in the tree that has exactly
one child. For example, Fig.~\ref{fig-simplification}E shows the
local trees along the genome for a simulated ARG
(Fig.~\ref{fig-simplification}A). Consider the right-most tree,
and the path above node $D$. Before $D$ meets its common ancestor
over the interval $(93, 100]$ with the other samples it passes through
nodes $H, K, N, R$ and $S$, reflecting the corresponding
traversal through the ARG (see section XXX). Such unary nodes
in a local tree are not usually considered significant,
because they provide very little information and are not
detectable by information from the tree. % weak
The nodes in a tree usually mark the points where samples find a
\emph{common} ancestor, rather than focusing on specific
ancestors. For most purposes, we would consider an edge
directly joining sample $D$ to the node $T$ to be more
useful and informative. Simplification is the process of removing
these uninformative and (for most computational purposes)
inefficient unary paths.

The first type of simplification that we can perform is to remove
graph topology that is invisible to the samples. The most well
known example of such topology is a so-called
``diamond''~\citep{rasmussen2014genome}
in which the two parent nodes of a recombination immediately
join again into a common ancestor (e.g. $K,N,O$ and $R$ in
Fig.~\ref{fig-simplification}A).
Unless we are specifically
interested in the recombination event or ancestral genomes,
there is no information in this topology and the diamond can b
e replaced by a single edge. More generally, any
subgraph which is singly-connected in both the leafward and
rootward direction (a ``super-diamond'') is non-identifiable and can be
replaced by a single edge. This definition includes the case
of a single node that has one inbound and one outbound edge.
Fig.~\ref{fig-simplification}B shows the result of this type of
graph topology simplification.

Simplifying away diamonds will remove many unary nodes from the
local trees, there can still be nodes that are unary in all
of the local trees. For example, node $H$ is one of the
parents of the recombinant node $D$ is still present in
Fig.~\ref{fig-simplification}B after diamond removal, but
it is unary in all of the local trees since it has an
in-degree of $1$.
Fig.~\ref{fig-simplification}C shows the resulting graph
after removing these nodes.

The remaining nodes are MRCAs of some subset of the samples
at \emph{some} positions along the genome. We still have
some unary nodes in the local trees, but these nodes will
correspond to a coalescence in at least one other
local tree. For example, node $E$ is unary in the first tree
of Fig.~\ref{fig-simplification}E, but is either binary
or ternary in the other trees (recall this is a Wright-Fisher
simulation).

The final level of simplification is to remove \emph{all}
unary nodes from the local trees (Fig.~\ref{fig-simplification}D
and E). [And continue]

@@

Due to the computational expense of sampling ``little ARGs'', state-of-the-art
Monte Carlo methods typically rely on cheaper, approximate models.
The most prominent examples are \texttt{ARGWeaver} \citep{rasmussen2014genome}
and \texttt{Arbores} \citep{heine2018bridging}. The former also utilises a time
discretisation approximation, but scales to dozens of mammal-like genomes.
The \texttt{ClonalOrigin} method \citep{didelot2010inference,
medina2020speeding} uses the \texttt{ClonalFrame}
model and data structure, though it is not as scalable as \texttt{ARGWeaver}.
